---
title: "Project 2 -- Regression Battleship"
author: "Jacob Jashinsky"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    code_folding: hide
---
```{r, warning=FALSE}
library(pander)
```
## Your Final Guess at My Model

Your final guess at my model.

<!-- Enter your model in math notation below. -->

$$
  Y_i = \beta_0  + \beta_1X_8 + \beta_2X_{15} + \beta_3X_{15}X_8 + \beta_4X_8X_6 + \epsilon_i 
$$

with your 95% confidence interval estimates of the coefficients as follows.

```{r}
sdat <- read.csv("sDat.csv", header=TRUE)

## EDIT THIS PART #################################################
## Input your lm(Y ~ ... ,data=sdat)
final.lm <- lm(Y ~ X8 + X15 + X15:X8 + X8:X6, data=sdat) #replace the "1" with your model
###################################################################

## This will create nice output for you.
mytable <- round(confint(final.lm, level=0.95), 2)
betas <- paste0("$\\beta_", 0:(length(final.lm$coefficients)-1), "$")
rownames(mytable) <- betas 
colnames(mytable) <- c("Lower", "Upper")
pander(mytable)
```

## Graph
```{r}
for (i in 1:50){
  if(sdat$X6[i]==1 & sdat$X15[i]==1){
  sdat$X21[i] <- 4
  }
  if(sdat$X6[i]==1 & sdat$X15[i]==0){
  sdat$X21[i] <- 3
  }
  if(sdat$X6[i]==0 & sdat$X15[i]==1){
  sdat$X21[i] <- 2
  }
  if(sdat$X6[i]==0 & sdat$X15[i]==0){
  sdat$X21[i] <- 1
  }
  
}
b <- final.lm$coefficients
plot(Y ~ X8, data=sdat, col=X21)
abline(b[1], b[2])
abline(b[1], b[2]+b[5], col="green")
abline(b[1] + b[3], b[2] + b[4] + b[5], col="blue")
abline(b[1] + b[3], b[2] + b[4], col="red")
```


<style>
#points {
  font-size:.8em;
  padding-left:5px;
  font-weight:bold; 
  color:#317eac;
}

#recpoints {
  font-size:.8em;
  padding-left:5px;
  font-weight:bold; 
  color:#7eac31;
}

#datalink {
  font-size:.5em;
  color:#317eac;
  padding-left:5px;
}

#headnote {
  font-size:.6em;
  color:#787878;
}

#note {
  font-size:.8em;
  color:#787878;
}

#headpoints {
 font-size:12pt;
 color: #585858; 
 padding-left: 15px;
}
</style>


<footer>
</footer>



 

 

 

 