---
title: "The Math 425 Notebook"
output: 
  html_document:
    theme: flatly
---



```{r setup, echo=FALSE}
library(tree)
pointSize = .8
pointSizeAccent = .4

pointColor = "gray41"
pointColor2 = "gray68"
pointColorAccent = "gray60"

lineColor = "gray58"
lineColor2 = "gray68"
```



## Unit 1

<a href=SimpleLinearRegression.html><div class="plotbox">

<center>
**Simple Linear**
</center>

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
par(mai=c(0,0,0,0))
set.seed(66)
x <- runif(50,0,20)
y <- 2+.8*x+rnorm(50,0,1.2)

plot(y ~ x, pch=16, cex=pointSize, xaxt='n', yaxt='n', xlab="", ylab="", col=pointColor, bty='n')
tlm <- lm(y ~ x)
abline(tlm, col=lineColor, lwd=2)
points(y ~ x, pch=16, cex=pointSizeAccent, col=pointColorAccent)

```

</div></a>




<a href=SimpleLinearRegressionRemediation.html><div class="plotbox">

<center>
**Remediation**
</center>

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
par(mai=c(0,0,0,0))
set.seed(656)
x <- sort(runif(50,0,20))
y <- 3 + 2.5*x + rnorm(50,0,seq(1,25, length.out=50))

plot(y ~ x, pch=16, cex=pointSize, xaxt='n', yaxt='n', xlab="", ylab="", col=pointColor, bty='n')
tlm <- lm(y ~ x)
abline(tlm, col=lineColor, lwd=2)
points(y ~ x, pch=16, cex=pointSizeAccent, col=pointColorAccent)
```

</div></a>

<a href=PolynomialRegression.html><div class="plotbox">

<center>
**Transformations**
</center>

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
par(mai=c(0,0,0,0))
set.seed(67)
x <- runif(50,10,20)
y <- log(1 + 250*x + rnorm(50,0,450))

plot(y ~ x, pch=16, cex=pointSize, xaxt='n', yaxt='n', xlab="", ylab="", col=pointColor, bty='n')
tlm <- lm(y ~ x + I(x^2) + I(x^3))
b <- coef(tlm)
curve(b[1]+b[2]*x+b[3]*x^2+b[4]*x^3, col=lineColor, lwd=2, add=TRUE)
points(y ~ x, pch=16, cex=pointSizeAccent, col=pointColorAccent)
```

</div></a>







<div style="clear:both;"></div>

<br/><br/>


## Unit 2

<div class="plotbox">

<center>
**Multiple Linear**
</center>

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
par(mai=c(0,0,0,0))
set.seed(667)
x <- runif(50,0,20)
x2 <- runif(50,0,20)
y <- 3+.8*x+rnorm(50,0,3.2)
y2 <- 4+2.3*x2+rnorm(50,0,3.2)

plot(y ~ x, pch=16, cex=pointSize, xaxt='n', yaxt='n', xlab="", ylab="", col=pointColor, bty='n', ylim=c(-10,55), xlim=c(0,20))
points(y2 ~ x2, pch=16, cex=pointSize,col=pointColor2)
mydat <- data.frame(y=c(y,y2), x=c(x,x2), g=rep(c(0,1),each=50))
tlm <- lm(y ~ x * g, data=mydat)
b <- coef(tlm)
abline(b[1],b[2], col=lineColor, lwd=2)
abline(b[1]+b[3],b[2]+b[4], col=lineColor2, lwd=2)
points(y ~ x, pch=16, cex=pointSizeAccent, col=pointColorAccent)
```

</div>

<div class="plotbox">

<center>
**High Dimensional**
</center>

```{r, include=FALSE}
library(plotly)
```

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
par(mai=c(0,0,0,0))
set.seed(66)
x <- runif(50,0,20)
y <- 2+.8*x+rnorm(50,0,1.2)

p <- plot_ly(x = ~x, y = ~x, z = ~y, color = rep(c("one","two"),25), colors = c('#BF382A', '#0C4B8E')) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'Weight'),
                     yaxis = list(title = 'Gross horsepower'),
                     zaxis = list(title = '1/4 mile time')))

```

</div>


<div class="plotbox">

<center>
**Model Selection**
</center>

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
thedata <- airquality[,-c(3,6)]
colnames(thedata) <- c("Y","X1","X2","X3")
pairs(thedata, pch=16, cex=pointSize, xaxt='n', yaxt='n', col=pointColor, bty='n', oma=c(0,0,0,0), gap=.25, cex.labels = .8)
```

</div>


<div style="clear:both;"></div>


<br/><br/>





## Unit 3

<div class="plotbox">

<center>
**Poisson**
</center>

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
par(mai=c(0,0,0,0))
set.seed(667)
x <- runif(50,0,20)
x2 <- runif(50,0,20)
y <- 3+.8*x+rnorm(50,0,3.2)
y2 <- 4+2.3*x2+rnorm(50,0,3.2)

plot(y ~ x, pch=16, cex=pointSize, xaxt='n', yaxt='n', xlab="", ylab="", col=pointColor, bty='n', ylim=c(-10,55), xlim=c(0,20))
points(y2 ~ x2, pch=16, cex=pointSize,col=pointColor2)
mydat <- data.frame(y=c(y,y2), x=c(x,x2), g=rep(c(0,1),each=50))
tlm <- lm(y ~ x * g, data=mydat)
b <- coef(tlm)
abline(b[1],b[2], col=lineColor, lwd=2)
abline(b[1]+b[3],b[2]+b[4], col=lineColor2, lwd=2)
points(y ~ x, pch=16, cex=pointSizeAccent, col=pointColorAccent)
```

</div>

<div class="plotbox">

<center>
**Trees**
</center>

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
par(mai=c(0,0,0,0))
tree.model <- tree(Species ~ Sepal.Width + Petal.Width, data=iris)
palette(c(pointColor2,pointColorAccent,pointColor))
plot(Sepal.Width ~ Petal.Width, data=iris, pch=19, col=Species, cex=pointSize)
partition.tree(tree.model, add=TRUE, col=rgb(1,1,1,0))

```

</div>


<div class="plotbox">

<center>
**Logistic**
</center>

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
par(mai=c(0,0,0,0))
set.seed(151)
x <- runif(30)
thedata <- data.frame(x=x,y=ifelse(x+rnorm(30,0,.1) > .5, 1, 0))
plot(y ~ x, data=thedata, pch=16, cex=pointSize, xaxt='n', yaxt='n', col=pointColor, bty='n',  ylim=c(-0.25,1.25))
myglm <- glm(y ~ x, data=thedata, family=binomial)
b <- coef(myglm)
curve(exp(b[1]+b[2]*x)/(1+exp(b[1]+b[2]*x)), add=TRUE, col=lineColor)
points(y ~ x, data=thedata, cex=pointSizeAccent, col=pointColorAccent, lwd=2)
```

</div>










<style>
a {
    color: #5a5a5a;
}

.plotbox {
    float: left;
    width: 175px;
    border: 1px solid #cccccc;
    background-color: #f8f8f8;
    border-radius: 3px;
    padding-left: 15px;
    padding-top: 5px;
    padding-bottom: 10px;
    padding-right: 10px;
    font-family: Monospace, sans-serif;
    margin-top: .5em;
}

.plotbox:hover {
    width: 175px;
    border: 1px solid #dddddd;
    background-color: #c1c1c1;
    border-radius: 3px;
    padding-left: 15px;
    padding-top: 5px;
    padding-bottom: 10px;
    padding-right: 10px;
    font-family: Monospace, sans-serif;
    color: #1a1a1a;
    margin-top: .5em;
}

</style>
