---
title: "Hard Work 9.2"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    code_folding: hide
---

```{r}
# This function will be useful later on.
makeTable9.3 <- function(thedata, nbest=1){
  
  if (nbest < 1 | nbest > dim(thedata)[2]-1){
    warning("nbest recoded to 1")
    nbest <- 1
  }
  
  cnames <- colnames(thedata)

  Yat <- grep("Y",cnames)
  
  if (length(Yat)>1)
    stop("Currently you have: ", paste(cnames[Yat], ", ", sep=""), 
               " in your data set.\n There can be only one response variable.\n",
               sep="")

  colnames(thedata)[Yat] <- "Y"
  Xat <- grep("X",cnames)
  
  require("leaps")
  require("qpcR")

  reg <- regsubsets(thedata[,Xat],thedata[,Yat], nbest=nbest, nvmax=length(cnames)-1)

  regs <- summary(reg)
  
  models <- sapply(1:length(row.names(regs$which)), function(i){
    colnames(regs$which)[which(regs$which[i,])][-1]
  })
  
  pressdata <<- thedata
  
  PRESS <- sapply(models, function(model){
    tmpf <- formula(paste("Y~",
                          paste(model,
                                c(rep("+",length(model)-1), ""), 
                                sep="", 
                                collapse=""), 
                          sep=""))
    tmp <- lm(tmpf, data=pressdata)
    tmp$call$formula <- eval(tmpf)
    PRESS(tmp, verbose=FALSE)$stat
  })
  
  n <- nrow(thedata)
  p <- as.numeric(row.names(regs$which))+1
 
  
  AIC <- regs$bic-log(n)*p-2*p

  
  AIC2 <- sapply(models, function(model){
    tmpf <- formula(paste("Y~",
                          paste(model,
                                c(rep("+",length(model)-1), ""), 
                                sep="", 
                                collapse=""), 
                          sep=""))
    tmp <- lm(tmpf, data=pressdata)
    tmp$call$formula <- eval(tmpf)
    AIC(tmp)
  })
  
  tmp <- data.frame(p=p, 
             model= sapply(1:length(models), function(i) {
                      paste(models[[i]], collapse=",")
                    }),
             SSEp=regs$rss,
             R2p=regs$rsq,
             R2ap=regs$adjr2,
             Cp=regs$cp,
             AICp=AIC2,
             SBCp=regs$bic,
             PRESSp=PRESS)
  
  rbind(data.frame(p=1,
                   model="(intercept)",
                   SSEp=regs$rss[1]/(1-regs$rsq[1]),
                   R2p=0,
                   R2ap=0,
                   Cp=NA,
                   AICp=NA,
                   SBCp=NA,
                   PRESSp=NA),tmp)
}
```


## Instructions

1. Study Section 9.6 -- "Model Validation"

2. Attempt and submit <span id=points style="padding-left:0px;">{15}</span> Hard Work Points by Saturday at 11:59 PM.    


## Reading Points <span id=headpoints>{7} Possible</span>

### Section 9.6 <span id=rrecpoints>{7}</span><span id=report>{ 7 / 7 }</span>


## Application Points <span id=headpoints>{8} Possible</span>


<a id=datalink style="font-size:.9em;" target="_blank" href=http://www.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/index.html>Data Files</a>

### Problem 9.21 <span id=recpoints>{2}</span><span id=report>{ 2 / 2 }</span> 

```{r}
# Code to read in the data
# It has to be merged from the 1.20 and 8.15 data files.
p9.10 <- read.table("http://www.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%209%20Data%20Sets/CH09PR10.txt", header=FALSE)
# Give it nice column names for conevience:
colnames(p9.10) <- c("Y","X1","X2","X3","X4")

# Show the first part of the data:
#head(p9.10)
```

The model is $Y_i = \beta_0 + \beta_1 X1 + \beta_2 X3 + \beta_3 X4$
```{r, warning=FALSE}
makeTable9.3(p9.10)
```

```{r}
summary(lm(Y ~ X1 + X3 + X4, data=p9.10))
```

Because PRESS_p is so over a 100 larger than SSE_p it is reasonable to suggest that the models predictive ability is lower than the model will claim, and that MSE_p is a biased estimator. 


### Problem 9.22 <span id=recpoints>{6}</span><span id=report>{ 6 / 6}</span> 

In part (b), swap each occurrence of the statement "Problem 9.18a" for "the significant model suggested by Problem 9.10c".

```{r}
# Code to read in the data
# It has to be merged from the 1.20 and 8.15 data files.
p9.22 <- read.table("http://www.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%209%20Data%20Sets/CH09PR22.txt", header=FALSE)
# Give it nice column names for conevience:
colnames(p9.22) <- c("Y","X1","X2","X3","X4")

# Show the first part of the data:
#head(p9.10)
```

**a.**


```{r}
cor(p9.10)
cor(p9.22)
```

The correlation matrix shows that the training sample is similar in correlation to the correlations of the prediction set.  

X2 is a little bit different, but considering that X2 is not in the model this won't be any concern for us. 

**b.**

```{r}
summary(lm(Y ~ X1 + X3 + X4, data=p9.10))
```

```{r}
summary(lm(Y ~ X1 + X3 + X4, data=p9.22))
```

The intercept has the most deviance, but all other coefficents and standard errors are very similar to the each other.  



```{r}
4.072^2
4.072*21
4.284^2
4.284*21
```


**c.**

MSPR = 4.284
MSE = 4.072

the difference is .21 

This may be enough deviations to call it into question, but I would suggest that it is reasonably close. 

This contradicts previous findings in p.21 which I concluded that MSE was baised. Because of the previous findings I am more likely to believe that MSE is baised. 

**d.** 

```{r}
total <- rbind(p9.10, p9.22)
```


```{r}
summary(lm(Y ~ X1 + X3 + X4, data=total))
```

All estimated standard deviations have been reduced when both data sets were combined. 



<style>
#points {
  font-size:.8em;
  padding-left:5px;
  font-weight:bold; 
  color:#317eac;
}

#recpoints {
  font-size:.8em;
  padding-left:5px;
  font-weight:bold; 
  color:#7eac31;
}

#rrecpoints {
  font-size:1em;
  padding-left:5px;
  font-weight:bold; 
  color:#7eac31;
}

#report {
  font-size:.7em;
  padding-left:15px;
  font-weight:normal; 
  color:#5a5a5a;
}

#datalink {
  font-size:.5em;
  color:#317eac;
  padding-left:5px;
}

#headnote {
  font-size:.6em;
  color:#787878;
}

#note {
  font-size:.8em;
  color:#787878;
}

#headpoints {
 font-size:12pt;
 color: #585858; 
 padding-left: 15px;
}
</style>



<footer>
</footer>



 

 

 

 